
---
title: "Machine Learning Project Part 1"
date: "Sunday, August 24, 2014"
---

##Project Background

There are many devices that measure activity of a person.  This study is using this data to predict the activity a person took part in.

##Data

The training data and test data are provided by: http://groupware.les.inf.puc-rio.br/har.  The training and test data are at the below links:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


##R code for creating prediction algorithm

First setting working directory

setwd("C:/Users/Brandon/Downloads")

##loading necessary packages for our analysis

```{r}
library(caret)
library(kernlab)
library(rpart)
```

##Loading and creating data frames for analysis

```{r}
train_file <- "pml-training.csv"
test_file <- "pml-test.csv"

train_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

if (!file.exists(test_file)) {
  download.file(test_url,test_file)
}
if (!file.exists(train_file)) {
  download.file(train_url,train_file)
}

training_data <-read.csv(train_file, header=TRUE,sep=",")
testing_data <-read.csv(test_file, header=TRUE,sep=",")
```

##Modifying training and testing files and subsetting the columns that we need

```{r}
removecols<-grep("avg|stddev|var|min|max|amplitude|skewness|kurtosis|timestamp|user_name|new_window",names(testing_data),value=FALSE)
testing_data<-testing_data[,-removecols]

removecols<-grep("avg|stddev|var|min|max|amplitude|skewness|kurtosis|timestamp|user_name|new_window",names(training_data),value=FALSE)
training_data<-training_data[,-removecols]

testing_data[,1]<-NULL
training_data[,1]<-NULL
```


##Subsetting and cleaning the data for run an algorithm just on the training dataset, not test data.  75% training data, 25% testing

train_data<-createDataPartition(y=training_data$classe,p=0.75,list=FALSE)
training_data_subset<-training_data[train_data,]
testing_data_subset<-training_data[-train_data,]

##Using principal components analysis what compenents have the most explaining power

```{r eval = FALSE}
pca_1 <-preProcess(training_data_subset[,-54],method="pca",pcaComp=10)
predict_train<-predict(pca_1,training_data_subset[,-54])
predict_test<-predict(pca_1,testing_data_subset[,-54])
```

##Building prediction model using random forests since all explanatory variables are continuous and outcome is a categorical field

##Also, looking into out of sample error between observed and predicted values & estimates

```{r eval = FALSE}
library(randomForest)

install.packages('e1071', dependencies=TRUE)

library(e1071)

random_forest <- train(x=predict_train,y=training_data_subset$classe,method="rf")

random_forest_fit <-predict(random_forest, predict_test)

confusionMatrix(random_forest_fit,predict_test$classe)
```










